---
title: 'Homework 1 - CUNY DATA 624 Sum I'
subtitle: 'Group 3: Daniel Sullivan, Jeff Parks, Lwin Shwe, Matthew Katz'
format:
  docx:
    highlight-style: arrow
    reference-doc: word-custom-theme-doc.docx
page-layout: article
editor: visual
---

```{r setup, include=FALSE}
# chunks
knitr::opts_chunk$set(echo=TRUE, eval=TRUE, include=TRUE, message=FALSE, 
warning=FALSE, fig.show="hold", out.width="100%", fig.height=3)

# libraries
library(tidyverse)
library(kableExtra)
library(fpp2)

# ggplot
theme_set(theme_light())
```

```{r functions, include=FALSE}
nice_table <- function(df, cap=NULL, cols=NULL, dig=3, fw=F){
  if (is.null(cols)) {c <- colnames(df)} else {c <- cols}
  table <- df %>% 
    kable(caption=cap, col.names=c, digits=dig) %>% 
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      html_font = 'monospace',
      full_width = fw)
  return(table)
}
```

## HA Exercise 2.1

**Use the help function to explore what the series `gold`, `woolyrnq` and `gas` represent.** a. **Use `autoplot()` to plot each of these in separate plots.**

b.  What is the frequency of each series? Hint: apply the `frequency()` function.

c.  Use `which.max()` to spot the outlier in the `gold` series. Which observation was it?

{{< pagebreak >}}

## HA Exercise 2.3

Download some monthly Australian retail data from the book website. These represent retail sales in various categories for different Australian states, and are stored in a MS-Excel file.

Select one of the time series ... explore your chosen retail time series using the following functions: autoplot(), ggseasonplot(), ggsubseriesplot(), gglagplot(), ggAcf()

Can you spot any seasonality, cyclicity and trend? What do you learn about the series?

{{< pagebreak >}}

## HA Exercise 6.2

The `plastics` data set consists of the monthly sales (in thousands) of product A for a plastics manufacturer for five years.

a.  Plot the time series of sales of product A. Can you identify seasonal fluctuations and/or a trend-cycle?

b.  Use a classical multiplicative decomposition to calculate the trend-cycle and seasonal indices.

c.  Do the results support the graphical interpretation from part a?

d.  Compute and plot the seasonally adjusted data.

e.  Change one observation to be an outlier (e.g., add 500 to one observation), and recompute the seasonally adjusted data. What is the effect of the outlier?

f.  Does it make any difference if the outlier is near the end rather than in the middle of the time series?
