---
title: 'Homework - CUNY DATA 624 Sum I'
subtitle: 'Group 3: Daniel Sullivan, Jeff Parks, Lwin Shwe, Matthew Katz'
format:
  docx:
    highlight-style: arrow
    reference-doc: word-custom-theme-doc.docx
page-layout: article
editor: visual
---

```{r setup, include=FALSE}
# chunks
knitr::opts_chunk$set(echo=TRUE, eval=TRUE, include=TRUE, message=FALSE, 
warning=FALSE, fig.show="hold", out.width="100%", fig.height=3)

# libraries
library(tidyverse)
library(kableExtra)
library(fpp2)

# ggplot
theme_set(theme_light())
```

## HA Exercise 2.1

**Use the help function to explore what the series `gold`, `woolyrnq` and `gas` represent.**

-   `gold` is a time-series dataset of daily morning gold prices in US dollars, from Jan 1, 1985 through Mar 31, 1989.
-   `woolyrng` is a time-series dataset of quarterly production of woolen yarn in Australia, in metric tons (tonnes), from Mar 1965 through Sep 1994.
-   `gas` is a time-series dataset of monthly gas production in Australia, units unknown, from 1956 through 1995.

------------------------------------------------------------------------

a.  **Use `autoplot()` to plot each of these in separate plots.**

```{r}
autoplot(gold) +
  ggtitle("Daily Morning Gold Prices") +
  xlab("Days") +
  ylab("US Dollars")
```

{{< pagebreak >}}

```{r}
autoplot(woolyrnq) +
  ggtitle("Quarterly Woollen Yarn Production: Australia") +
  xlab("Year") +
  ylab("Metric Tons")
```

```{r}
autoplot(gas) +
  ggtitle("Monthly Gas Production: Australia") +
  xlab("Years") +
  ylab("Units")
```

b.  **What is the frequency of each series? Hint: apply the `frequency()` function.**
    -   The frequency of `gold` is `r frequency(gold)` (daily)
    -   The frequency of `woolyrnq` is `r frequency(woolyrnq)` (quarterly)
    -   The frequency of `gas` is `r frequency(gas)` (yearly)\
c.  **Use `which.max()` to spot the outlier in the `gold` series. Which observation was it?**
    -   The maximum value in the `gold` series is observation **`r which.max(gold)`** with a value of **`r gold[770]`**.
